---
title: "Example of E-M with Gaussian mixture model"
author: "Rebecca"
date: "March 29, 2017"
output: pdf_document
---

We're going to practice applying the EM algorithm to a multivariate normal distribution where we are assuming there are $k = 3$ clusters. 

We're going to do so in two stages:

\begin{enumerate}
\item Get the basics down with a few iterations where we print the results
\item Embed into a \texttt{while} loop that will stop at the definition of convergence we choose: differences between the old and newer values of parameters get very small
\end{enumerate}

We're going to practice with example data on characteristics of wine from different 'cultivars' in Italy from the \texttt{gclus} package.


```{r}
# Load data
library(gclus)
library(ggplot2)
##3 clusters, 13 dimensions
##see if can implement
data(wine, package = "gclus")
head(wine)

##plot two dimensions and coor by class
ggplot(wine, aes(x = Alcohol, 
  y = Phenols, color = as.factor(Class))) +
  geom_point() 

##recover latent cluster labels- and then see how
##corresponds


```


\subsection{Step two: implement the steps described on Lecture slides 38 and 39}

\subsubsection{Step one: Initialize the parameters}

Set k to 3 

```{r}

##restrict wine data to Alcohol, Phenols, and class
library(dplyr)
wine2 <- wine %>%
        select(Class, Alcohol, Phenols)
head(wine2)

#three probability of cluster assignments
pi_init <- c(0.3333, 0.3333, 0.33333)

##divide data into thirds
cut1 <- round(nrow(wine2)/3)
cut2 <- round(cut1 + nrow(wine2)/3)


##initialize three 2-length mean vector
mu_1_init <- c(mean(wine2$Alcohol[1:cut1]),
               mean(wine2$Phenols[1:cut1]))
mu_2_init <- c(mean(wine2$Alcohol[cut1:cut2]),
               mean(wine2$Phenols[cut1:cut2]))
mu_3_init <- c(mean(wine2$Alcohol[cut2:nrow(wine2)]),
               mean(wine2$Phenols[cut2:nrow(wine2)])) 

##initialize sigma to identity matrix
sigma_1_init <- sigma_2_init <- sigma_3_init <- matrix(c(1, 0, 0, 1),
                                                       nrow = 2, ncol = 2,
                                                       byrow= TRUE)
sigma_1_init



```


\subsubsection{Step two: compute responsibilities for each observation and cluster}


First, let's try with observation 1 and observation 2 and then see how vector of responsibilities corresponds to observed classes

```{r}
##store the denominator
denom_resp_1 <- pi_init[1] * dmvnorm(x = c(wine2[1, "Alcohol"],
                      wine2[1, "Phenols"]),
                      mean = mu_1_init,
                      sigma = sigma_1_init) +
        pi_init[2] * dmvnorm(x = c(wine2[1, "Alcohol"],
                      wine2[1, "Phenols"]),
                      mean = mu_2_init,
                      sigma = sigma_2_init) +
        pi_init[3] * dmvnorm(x = c(wine2[1, "Alcohol"],
                      wine2[1, "Phenols"]),
                            mean = mu_3_init,
                          sigma = sigma_3_init)


r_11 <- pi_init[1] * dmvnorm(x = c(wine2[1, "Alcohol"],
                    wine2[1, "Phenols"]),
                    mean = mu_1_init,
                    sigma = sigma_1_init)/denom_resp_1    
r_11
   

r_12 <- pi_init[2] * dmvnorm(x = c(wine2[1, "Alcohol"],
                            wine2[1, "Phenols"]),
                    mean = mu_2_init,
                    sigma = sigma_2_init)/denom_resp_1   
r_12

r_13 <- pi_init[3] * dmvnorm(x = c(wine2[1, "Alcohol"],
                            wine2[1, "Phenols"]),
                    mean = mu_3_init,
                    sigma = sigma_3_init)/denom_resp_1   
r_13

##try with second data point
denom_resp_2 <- pi_init[1] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                      wine2[nrow(wine2), "Phenols"]),
                      mean = mu_1_init,
                      sigma = sigma_1_init) +
        pi_init[2] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                      wine2[nrow(wine2), "Phenols"]),
                      mean = mu_2_init,
                      sigma = sigma_2_init) +
        pi_init[3] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                      wine2[nrow(wine2), "Phenols"]),
                            mean = mu_3_init,
                          sigma = sigma_3_init)

##compute responsibilities
r_last1 <- pi_init[1] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                            wine2[nrow(wine2), "Phenols"]),
                    mean = mu_1_init,
                    sigma = sigma_1_init)/denom_resp_2    
r_last1
   
r_last2 <- pi_init[2] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                            wine2[nrow(wine2), "Phenols"]),
                    mean = mu_2_init,
                    sigma = sigma_2_init)/denom_resp_2   
r_last2

r_last3 <- pi_init[3] * dmvnorm(x = c(wine2[nrow(wine2), "Alcohol"],
                            wine2[nrow(wine2), "Phenols"]),
                    mean = mu_3_init,
                    sigma = sigma_3_init)/denom_resp_2   
r_last3

##look at each compared to observed
wine2[1, "Class"]; c(r_11, r_12, r_13)

#
wine2[nrow(wine2), "Class"]; c(r_last1, r_last2, r_last3)

##compute responsibilities for all n observations
store_r <- matrix(NA, nrow = nrow(wine2), ncol = 3)
denom_resp <- rep(NA, nrow(wine2))
for(i in 1:nrow(wine2)){
  #denominator for responsibilities
  denom_resp[i] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_1_init,
                      sigma = sigma_1_init) +
        pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_2_init,
                      sigma = sigma_2_init) +
        pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                            mean = mu_3_init,
                          sigma = sigma_3_init)
  
  #responsibility 1
  store_r[i, 1] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_1_init,
                    sigma = sigma_1_init)/denom_resp[i]
  
  store_r[i, 2] <- pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_2_init,
                    sigma = sigma_2_init)/denom_resp[i]
  
  store_r[i, 3] <- pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_3_init,
                    sigma = sigma_3_init)/denom_resp[i]
  
}
head(store_r)
summary(store_r)



```

\subsubsection{Step three: m-step}


First, update the probabilities of belonging to each cluster

```{r}
pi_new <- colSums(store_r)/nrow(store_r)
pi_new

```

Then, update the mean for each cluster as weighted average

```{r}
n_1 <- sum(store_r[, 1])
n_2 <- sum(store_r[, 2])
n_3 <- sum(store_r[, 3]) 
mu_1_new <- c(sum(store_r[, 1] * wine2[, "Alcohol"])/n_1,
              sum(store_r[, 1] * wine2[, "Phenols"]/n_1)) 
mu_1_new; mu_1_init
mu_2_new <- c(sum(store_r[, 2] * wine2[, "Alcohol"])/n_2,
              sum(store_r[, 2] * wine2[, "Phenols"]/n_2)) 
mu_2_new; mu_2_init
mu_3_new <- c(sum(store_r[, 3] * wine2[, "Alcohol"])/n_3,
              sum(store_r[, 3] * wine2[, "Phenols"]/n_3)) 
mu_3_new; mu_3_init

```

Finally, with those new mu's, compute updated sigma for each cluster

```{r}
mat_function <- function(i, mu, k, allmat){
  x_min_mu <- rbind(wine2[i, c("Alcohol")] - mu[1],
              wine2[i, c("Phenols")] - mu[2])
  x_min_mu_t <- t(x_min_mu)
  prod_x_min_mu <- x_min_mu %*% x_min_mu_t
  r_prod_x_min_mu <- store_r[i, k] * prod_x_min_mu
  all_mat[] <- r_prod_x_min_mu
}

##store results- make into a matrix
sigma_1_all <- sapply(1:178, mat_function, mu = mu_1_new,
               k = 1, allmat = list())
sigma_1_mat <- matrix(c(sum(sigma_1_all[1, ]),
                        sum(sigma_1_all[2, ]),
                        sum(sigma_1_all[3, ]),
                        sum(sigma_1_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
#rescale by n1
sigma_1_new <- (1/n_1) * sigma_1_mat
sigma_1_new; sigma_1_init

##repeat for sigma_2 and sigma_3
sigma_2_all <- sapply(1:178, mat_function, mu = mu_2_new,
               k = 2, allmat = list())
sigma_2_mat <- matrix(c(sum(sigma_2_all[1, ]),
                        sum(sigma_2_all[2, ]),
                        sum(sigma_2_all[3, ]),
                        sum(sigma_2_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
#rescale by n
sigma_2_new <- (1/n_2) * sigma_2_mat
sigma_2_new; sigma_2_init

##sigma_3
sigma_3_all <- sapply(1:178, mat_function, mu = mu_3_new,
               k = 3, allmat = list())
sigma_3_mat <- matrix(c(sum(sigma_3_all[1, ]),
                        sum(sigma_3_all[2, ]),
                        sum(sigma_3_all[3, ]),
                        sum(sigma_3_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
#rescale by n
sigma_3_new <- (1/n_3) * sigma_3_mat
sigma_3_new; sigma_3_init

```

Then, feed those new mu and sigma back into responsibilities step and repeat

Practice with one update of responsibilities

```{r}
##assign new values to initial values
pi_init <- pi_new
mu_1_init <- mu_1_new
mu_2_init <- mu_2_new
mu_3_init <- mu_3_new
sigma_1_init <- sigma_1_new
sigma_2_init <- sigma_2_new
sigma_3_init <- sigma_3_new

store_r_2 <- matrix(NA, nrow = nrow(wine2), ncol = 3)
denom_resp <- rep(NA, nrow(wine2))
for(i in 1:nrow(wine2)){
  #denominator for responsibilities
  denom_resp[i] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_1_init,
                      sigma = sigma_1_init) +
        pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_2_init,
                      sigma = sigma_2_init) +
        pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                            mean = mu_3_init,
                          sigma = sigma_3_init)
  
  #responsibility 1
  store_r_2[i, 1] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_1_init,
                    sigma = sigma_1_init)/denom_resp[i]
  
  store_r_2[i, 2] <- pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_2_init,
                    sigma = sigma_2_init)/denom_resp[i]
  
  store_r_2[i, 3] <- pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_3_init,
                    sigma = sigma_3_init)/denom_resp[i]
  
}
head(store_r_2)
summary(store_r_2)

```


Create two contrasting plots- one with real versus assigned to highest from iteration 1 and another with real versus assigned to highest with iteration 2

```{r}
#assign each observation to their m
wine3 <- wine2 %>%
          mutate(label_iter1 = max.col(store_r, ties.method = 'first'),
                 label_iter2 = max.col(store_r_2, ties.method = 'first')) 

wine3


```

Generalize into more iterations


```{r}
#for simplicity, define difference
#as between mu1 and mu2
pi_init <- c(0.333, 0.3333, 0.3333)
diff <- mean(abs(pi_init - pi_new))
diff
while(diff > 0.0001){
  
  #e-step of responsibiliites matrix
  store_r <- matrix(NA, nrow = nrow(wine2), ncol = 3)
  denom_resp <- rep(NA, nrow(wine2))
  for(i in 1:nrow(wine2)){
  #denominator for responsibilities
  denom_resp[i] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_1_init,
                      sigma = sigma_1_init) +
        pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                      mean = mu_2_init,
                      sigma = sigma_2_init) +
        pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                      wine2[i, "Phenols"]),
                            mean = mu_3_init,
                          sigma = sigma_3_init)
  
  #responsibility 1
  store_r[i, 1] <- pi_init[1] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_1_init,
                    sigma = sigma_1_init)/denom_resp[i]
  
  store_r[i, 2] <- pi_init[2] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_2_init,
                    sigma = sigma_2_init)/denom_resp[i]
  
  store_r[i, 3] <- pi_init[3] * dmvnorm(x = c(wine2[i, "Alcohol"],
                            wine2[i, "Phenols"]),
                    mean = mu_3_init,
                    sigma = sigma_3_init)/denom_resp[i]
  

  }
  
  ##m-step - update each parameter
  pi_new <- colSums(store_r)/nrow(store_r)

  #update n's 
  n_1 <- sum(store_r[, 1])
  n_2 <- sum(store_r[, 2])
  n_3 <- sum(store_r[, 3]) 
  
  #update mu's
  mu_1_new <- c(sum(store_r[, 1] * wine2[, "Alcohol"])/n_1,
              sum(store_r[, 1] * wine2[, "Phenols"]/n_1)) 
  mu_2_new <- c(sum(store_r[, 2] * wine2[, "Alcohol"])/n_2,
              sum(store_r[, 2] * wine2[, "Phenols"]/n_2)) 
  mu_3_new <- c(sum(store_r[, 3] * wine2[, "Alcohol"])/n_3,
              sum(store_r[, 3] * wine2[, "Phenols"]/n_3)) 

  #update sigma's
  sigma_1_all <- sapply(1:178, mat_function, mu = mu_1_new,
               k = 1, allmat = list())
  sigma_1_mat <- matrix(c(sum(sigma_1_all[1, ]),
                        sum(sigma_1_all[2, ]),
                        sum(sigma_1_all[3, ]),
                        sum(sigma_1_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
  #rescale by n1
  sigma_1_new <- (1/n_1) * sigma_1_mat

  ##repeat for sigma_2 and sigma_3
  sigma_2_all <- sapply(1:178, mat_function, mu = mu_2_new,
               k = 2, allmat = list())
  sigma_2_mat <- matrix(c(sum(sigma_2_all[1, ]),
                        sum(sigma_2_all[2, ]),
                        sum(sigma_2_all[3, ]),
                        sum(sigma_2_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
  #rescale by n
  sigma_2_new <- (1/n_2) * sigma_2_mat

  ##sigma_3
  sigma_3_all <- sapply(1:178, mat_function, mu = mu_3_new,
               k = 3, allmat = list())
  sigma_3_mat <- matrix(c(sum(sigma_3_all[1, ]),
                        sum(sigma_3_all[2, ]),
                        sum(sigma_3_all[3, ]),
                        sum(sigma_3_all[4, ])),
                        nrow = 2, ncol = 2, byrow = TRUE) 
  #rescale by n
  sigma_3_new <- (1/n_3) * sigma_3_mat
  
  ##before reassigning, calculate new diff
  diff <- mean(abs(pi_init - pi_new))
  
  #reassign new values to initial values
  ##assign new values to initial values
  pi_init <- pi_new
  mu_1_init <- mu_1_new
  mu_2_init <- mu_2_new
  mu_3_init <- mu_3_new
  sigma_1_init <- sigma_1_new
  sigma_2_init <- sigma_2_new
  sigma_3_init <- sigma_3_new
  
  
}

#view results
head(store_r)

#can then compare new means to actual means (we gave it really informative
#starting means to make sure responsibility step worked but does especially for 
#cluster 3 get closer- to start with less informative initial values, can reshuffle data
#so that it's no longer ordered by actual cluster)

```